---
title: "PA1 Analysis"
author: "Malcolm Neo"
date: "25 June, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Activity Monitoring Case Studies

## 1. Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## 2. Loading of Data

Following R code can be used to load the data necessary for the case study.

```{r load data}
# Read dataset from csv file
activity<-read.csv('C:/Users/User/Documents/R/RepData_PeerAssessment1/activity.csv')
```

## 3. Analysis

### 3.1. Loading and Preprocessing of Data

The steps are taken to process the data:

1. Aggregate steps taken each day
2. Name the columns of the results

```{r echo=TRUE}
# Total number of steps taken each day
sum_step_by_day = aggregate(activity$steps, by = list(activity$date), sum,na.rm=TRUE)

names(sum_step_by_day) <- c("Date", "steps")

```

### 3.2. What is mean total number of steps taken per day?
#### 3.2.2. Histogram of the total number of steps taken each day

```{r TotalNumber, echo=TRUE}
# Histogram of the total number of steps taken each day
hist(sum_step_by_day$steps)
```

#### 3.2.3 Mean and Medium of steps taken each day

##### 3.2.3.1. Mean of steps taken each day
```{r MeanNumber, echo=TRUE}
mean(sum_step_by_day$steps)
```

##### 3.2.3.2. Medium of steps taken each day
```{r MedianNumber, echo=TRUE}
median(sum_step_by_day$steps)
```

### 3.3. What is the average daily activity pattern?
#### 3.3.1. Time series plot of the average number of steps taken
```{r TimeSeriesPlot, echo=TRUE}
steps_Interval<-aggregate(steps~interval,data=activity,mean,na.rm=TRUE)
plot(steps~interval,data=steps_Interval,type="l")
```

#### 3.3.2. The 5-minute interval that, on average, contains the maximum number of steps
```{r IntervalMaxStep, echo=TRUE}
steps_Interval[which.max(steps_Interval$steps),]$interval
```

### 3.4. Inputing missing values
```{r PopMissingData, echo=TRUE}

calStep<-function(interval){
    steps_Interval[steps_Interval$interval==interval,]$steps
}

# Filling the missing values in the dataset
activity_populated<-activity

count=0
for(i in 1:nrow(activity_populated)){
  if(is.na(activity_populated[i,]$steps)){
    activity_populated[i,]$steps<-calStep(activity_populated[i,]$interval)
    count=count+1
  }
}

# Sum of populated dataset
sum_step_by_day2<-aggregate(steps~date,data=activity_populated,sum)
```

#### 3.4.1. Histogram of the total number of steps taken each day after missing values are inputed
```{r TotalNumber2, echo=TRUE}

# Sum of populated dataset
sum_step_by_day2<-aggregate(steps~date,data=activity_populated,sum)

# Histogram, Mean and Median of new dataset
hist(sum_step_by_day2$steps)
```


#### 3.4.2. Mean and Medium of steps taken each day
##### 3.4.2.1. Mean of steps taken each day
```{r MeanNumber2, echo=TRUE}
mean(sum_step_by_day2$steps)
```

##### 3.4.2.2. Medium of steps taken each day
```{r MedianNumber2, echo=TRUE}
median(sum_step_by_day2$steps)
```

The mean value have increased due to the changes in the data.
The Frequency have changed due to the changes in the data.

### 3.5. Are there differences in activity patterns between weekdays and weekends?

```{r DayOfWeekPlot, echo=TRUE}
# Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
activity_populated$day=ifelse(as.POSIXlt(as.Date(activity_populated$date))$wday%%6==0,
                          "weekend","weekday")
activity_populated$day=factor(activity_populated$day,levels=c("weekday","weekend"))

steps_Interval2=aggregate(steps~interval+day,activity_populated,mean)

library(lattice)
#xyplot(steps~interval|factor(day),data=steps_Interval2,aspect=1/2,type="l")

xyplot(steps~interval | day, data = steps_Interval2,
       type = 'l',
       xlab = 'Interval',
       ylab = 'Number of Steps',
       layout = c(1,2))
```